---
title: "Exam 2 Notes"
author: "Anna OConnell"
date: "2026-03-05"
output: html_document
---

 Everything important we learn between exams 1 and 2. Note: exam is cumulative
 
## Multiple Linear Regression
 
 - when interpreting, keep all other predictors constant
 - need to engineer all variables so theyre all linear
 
 
```{r}
# read data in
mvdata <- read.csv('C:/Users/annae/Downloads/ISA391/MarketValueData.csv')
head(mvdata)

#Simple linear regression would just have only one predictor, no +
fit.mv <- lm(Mkt_Value_1000s~Sqr_Ft+House_Age, data=mvdata)
summary(fit.mv)
```


# Assumptions 
 
 **Linearity** each predictor plotted with y, is scatterplot linear? can also plot line of best fit \
 plot(college$Cost4_A, 
     college$md_earn_wne_p10,
     main= "Scatterplot of Cost vs Earning",
     xlab = "Cost",
     ylab = "Earn",
     col= "pink") \
abline(college.reg) \
 
 **Assumption 1: Mean e Zero** test that the mean of the e distribution is zero \
 mean(college.res) \

 **Assumption 2: Equal Variance** residual plot \
names(college.reg) \
college.fitted <- college.reg$fitted.values \
plot(college.fitted, college.res, ylab="residuals", xlab = "predicted earnings", main = "Residual Plot") \
abline(0,0)\

 
 **Assumption 3: Normality** histogram of e, qq normal quantile plot \
 wine.resid <- resid(wine.reg) \
qqnorm(wine.resid) \
qqline(wine.resid) \

hist(college.res) #gets histogram of residuals \

 **Assumption 4: Independence** is sample randomly selected? if so, residuals will be independent from one another. \
 
 why do i need only one plot for residuals but one for each predictor for linearity? because even for many predictors, each observation only has one residual 
 
 
# Code #

Scatterplot matrix: 
```{r}
pairs(mvdata) #scatterplot matrix to check lienarity 

pairs(~Mkt_Value_1000s+Sqr_Ft+House_Age, data=mvdata) # ~ means "by" want them all by each other
```

To get just coefficents table: 
```{r}
fit.mv <- lm(Mkt_Value_1000s~Sqr_Ft+House_Age, data=mvdata)
summary(fit.mv)$coefficients 
```

To get anova table: (test meaningful/not)
```{r}
anova(fit.mv)
```

## Interpretations

**Coefficents** 

**B0** when the size AND age of a home is zero, the estimated market value is $197.331. Not practical in this case because a house has to have a size and age > 0 \
   in this case, we can check summary(mvdata$predictor) and then check the minimum to see if it could be zero, avoid extrapolation \
**B1** when a house has one more square foot, the price will increase by $40.90 KEEPING THE HOUSE AGE CONSTANT \
**B2** when a house is one more year old, the price will decrease by $825.00 KEEPING THE SQUARE FOOTAGE CONSTANT\

## Overall Significance of Regression

**Hypothesis test**
**Ho** B1=B2=...=0 ; The regression is not meaningful
**Ha** At least one B is not 0; The regression is meaningful

**Degree of Freedom**
for model --> k: number of predictors
for error --> n-(k+1)
total --> n-1

```{r}
# pf() gives you everything to the left of the F stat, so p value is 1-pf
# and do pf(x,x,x, lower.tail=FALSE)
```
Now, if we want to test for an individual predictor, partial t-test \ 
Use Summary , coefficient for each predictor 

**Interpretations**
if p-value < alpha: predictor is useful / significant contribution to regression 
if p-value > alpha: with other predictors already in regression, adding predictor does not provide significant contribution

Get a P-value based on t test stat 

```{r}
pt(6.109, 39, lower.tail=FALSE)*2
```

**Confidence Intervals**

```{r}
confint(fit.mv, level=0.95)
#With same house age/with house age fixed, then if you increase one sqr foot of house, house price will increase by (x,x)
```

R^2 = proportion of observed variation in y

```{r}
predict(fit.mv,
        newdata=data.frame(House_Age=25, Sqr_Ft=1500),
        interval = 'confidence',
        level = 0.95)
```









